---
categories:  
- ""    #the front matter should be like the one found in, e.g., blog2.md. It cannot be like the normal Rmd we used
- ""
date: "2022-09-18"
description: Analysis of Movie Ratings Using IMDB Data # the title that will show up once someone gets to this page
draft: false
image: "" # save picture in \static\img\blogs. Acceptable formats= jpg, jpeg, or png . Your iPhone pics wont work

keywords: ""
slug: project1 # slug is the shorthand URL address... no spaces plz
title: Analysis of Movie Ratings Using IMDB Data
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
```

# Analysis of movies- IMDB dataset

We will look at a subset sample of movies, taken from the [Kaggle IMDB 5000 movie dataset](https://www.kaggle.com/carolzhangdc/imdb-5000-movie-dataset)

```{r,load_movies, warning=FALSE, message=FALSE}

movies <- read_csv(here::here("movies.csv"))
glimpse(movies)
skim(movies)

```

Besides the obvious variables of `title`, `genre`, `director`, `year`, and `duration`, the rest of the variables are as follows:

-   `gross` : The gross earnings in the US box office, not adjusted for inflation
-   `budget`: The movie's budget
-   `cast_facebook_likes`: the number of facebook likes cast members received
-   `votes`: the number of people who voted for (or rated) the movie in IMDB
-   `reviews`: the number of reviews for that movie
-   `rating`: IMDB average rating

## Use your data import, inspection, and cleaning skills to answer the following:

1.  Are there any missing values (NAs)? Are all entries distinct or are there duplicate entries?

    \- There is no missing values, if we use is.na(movies) to check. There are 54 duplicate entries.

2.  Produce a table with the count of movies by genre, ranked in descending order

    ```{r, count_of_movies_by_genre}
    count_of_movies_by_genre <- movies %>% 
      group_by(genre) %>% 
      count(genre) %>% 
      arrange(desc(n))
    count_of_movies_by_genre

    ```

3.  Produce a table with the average gross earning and budget (`gross` and `budget`) by genre. Calculate a variable `return_on_budget` which shows how many \$ did a movie make at the box office for each \$ of its budget. Ranked genres by this `return_on_budget` in descending order

    ```{r, avg_gross_budget_by_genre}
    # producing the new table
    avg_gross_budget_by_genre <- movies %>% 
      group_by(genre) %>% 
      summarise(avg_gross = mean(gross), avg_budget = mean(budget))

    # calculating return_on_budget
    return_on_budget <- avg_gross_budget_by_genre %>% 
      mutate(return_on_budget = avg_gross/avg_budget) %>% 
      select(return_on_budget) %>% 
      arrange(desc(return_on_budget))
    ```

4.  Produce a table that shows the top 15 directors who have created the highest gross revenue in the box office. Don't just show the total gross amount, but also the mean, median, and standard deviation per director.

    ```{r}
    movies %>% 
      group_by(director) %>% 
      summarise(gross_rev = sum(gross), mean_rev = mean(gross), median_rev = median(gross), SD_rev = sd(gross)) %>% 
      slice_max(n=15,order_by=gross_rev)
    ```

5.  Finally, ratings. Produce a table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed.

    ```{r, ratings}
    ratings_table <- movies %>% 
      group_by(genre) %>% 
      summarise(mean=mean(rating), min=min(rating), max=max(rating), median=median(rating),SD=sd(rating))

    # density graph of rating distribution
    ggplot(movies) +
      aes(x=rating) +
      geom_density() +
      theme_bw(base_size = 18) +
      labs(title = "Rating Distribution",
           x = NULL, 
           y = NULL, 
           caption = "Source: Kaggle IMDB 5000 movie dataset")
    ```

## Use `ggplot` to answer the following

-   Examine the relationship between `gross` and `cast_facebook_likes`. Produce a scatterplot and write one sentence discussing whether the number of facebook likes that the cast has received is likely to be a good predictor of how much money a movie will make at the box office. What variable are you going to map to the Y- and X- axes?

```{r, gross_on_fblikes}

ggplot(movies) +
  aes(x = cast_facebook_likes, y = gross) +
  geom_point(size=3) +
  geom_smooth(method='lm')
  labs (x = "Number of Likes for Cast on Faceboook", y = "Gross Earnings in the US Box Office") +
  theme_classic(base_size = 18) +
  scale_x_log10()

```

-   Examine the relationship between `gross` and `budget`. Produce a scatterplot and write one sentence discussing whether budget is likely to be a good predictor of how much money a movie will make at the box office.

```{r, gross_on_budget}
ggplot(movies) + 
  aes(x=budget, y=gross) + 
    geom_point(size =3) + 
    geom_smooth(method='lm', se=FALSE)+ 
  scale_x_log10() + 
    labs(title='relationship between budget and gross') +
  theme_bw(base_size=18)
```

-   Examine the relationship between `gross` and `rating`. Produce a scatterplot, faceted by `genre` and discuss whether IMDB ratings are likely to be a good predictor of how much money a movie will make at the box office. Is there anything strange in this dataset?

```{r, gross_on_rating}
ggplot(movies) + 
  aes(x=rating, y=gross, colour = genre) +
  geom_point(size=3) +
  facet_wrap(~genre) +
  labs(title = 'Relationship between Rating and Gross for each genre') +
  theme_bw(base_size=18) 
```
