---
categories:  
- ""    #the front matter should be like the one found in, e.g., blog2.md. It cannot be like the normal Rmd we used
- ""
date: "2022-09-18"
description: Analysis of Movie Ratings Using IMDB Data # the title that will show up once someone gets to this page
draft: false
image: "" # save picture in \static\img\blogs. Acceptable formats= jpg, jpeg, or png . Your iPhone pics wont work

keywords: ""
slug: project1 # slug is the shorthand URL address... no spaces plz
title: Analysis of Movie Ratings Using IMDB Data
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=9.25, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(infer)
```

## Summary of the Project

1.  Import and inspect the IMDB dataset using skimr and dplyr package functions.

2.  Clean the data set and conduct exploratory data analysis (EDA).

3.  Use ggplot2 package functions to visualize relationships between variables of interest. We want to investigate the relationship between gross earnings in the US box office [gross] and the number of facebook likes cast members received [likes], IMDB average rating [ratings] and the movie's budget [budget] respectively.

4.  An investigation into whether the mean IMDB ratings for movie director Steven Spielberg and Tim Burton are the same or not.

## Importing and Inspecting the IMDB Dataset

We looked at a subset sample of movies, taken from the [Kaggle IMDB 5000 movie dataset](https://www.kaggle.com/carolzhangdc/imdb-5000-movie-dataset)

```{r,load_movies, warning=FALSE, message=FALSE}

movies <- read_csv(here::here("movies.csv"))
glimpse(movies)
skim(movies)

```

Besides the obvious variables of `title`, `genre`, `director`, `year`, and `duration`, the rest of the variables are as follows:

-   `gross` : The gross earnings in the US box office, not adjusted for inflation
-   `budget`: The movie's budget
-   `cast_facebook_likes`: the number of facebook likes cast members received
-   `votes`: the number of people who voted for (or rated) the movie in IMDB
-   `reviews`: the number of reviews for that movie
-   `rating`: IMDB average rating

## Data Cleaning

1.  Firstly, check for missing values and duplicated entries. There is no missing values, if we use is.na(movies) to check. There are 54 duplicate entries.

2.  Make a table with the count of movies by genre, ranked in descending order

    ```{r, count_of_movies_by_genre}
    count_of_movies_by_genre <- movies %>% 
      group_by(genre) %>% 
      count(genre) %>% 
      arrange(desc(n))
    count_of_movies_by_genre

    ```

3.  Produce a table with the average gross earning and budget (`gross` and `budget`) by genre. Calculate a variable `return_on_budget` which shows how many \$ did a movie make at the box office for each \$ of its budget. Ranked genres by this `return_on_budget` in descending order

    ```{r, avg_gross_budget_by_genre}
    # producing the new table
    avg_gross_budget_by_genre <- movies %>% 
      group_by(genre) %>% 
      summarise(avg_gross = mean(gross), avg_budget = mean(budget))

    # calculating return_on_budget
    return_on_budget <- avg_gross_budget_by_genre %>% 
      mutate(return_on_budget = avg_gross/avg_budget) %>% 
      select(genre, return_on_budget) %>% 
      arrange(desc(return_on_budget))

    ```

4.  Produce a table that shows the top 15 directors who have created the highest gross revenue in the box office. Showing the total gross amount, as well as the mean, median, and standard deviation per director.

    ```{r}
    movies %>% 
      group_by(director) %>% 
      summarise(gross_rev = sum(gross), mean_rev = mean(gross), median_rev = median(gross), SD_rev = sd(gross)) %>% 
      slice_max(n=15,order_by=gross_rev)
    ```

5.  Finally, ratings. Produce a table that describes how ratings are distributed by genre. Find the mean, min, max, median, SD and draw a graph to show how ratings are distributed.

    ```{r, ratings}
    ratings_table <- movies %>% 
      group_by(genre) %>% 
      summarise(mean=mean(rating), min=min(rating), max=max(rating), median=median(rating),SD=sd(rating))

    # density graph of rating distribution
    ggplot(movies) +
      aes(x=rating, colour = genre) +
      geom_density(size = 1.3) +
      facet_wrap(~genre) +
      theme_bw(base_size = 18) +
      labs(title = "Distribution of IMDB Ratings by Film Genre ",
           x = NULL, 
           y = NULL, 
           caption = "Source: Kaggle IMDB 5000 Movie Dataset") +
       theme(legend.position = "NONE")
    ```

## Examine the Relationship of Variables of Interest

-   Examine the relationship between `gross` and `cast_facebook_likes`.
-   The scatter plot below shows that there is a slight positive correlation between 'gross' and 'cast_facebook_like', although there exist many outliers, such as when likes is less than 100. Therefore, the cast is not likely to be a good predictor. However, many data points lies outside the a line of best fit, therefore, we cannot say with good confidence that it is a good predictor.

```{r, gross_on_fblikes}

ggplot(movies) +
  aes(x = cast_facebook_likes, y = gross) +
  geom_point(size=3) +
  geom_smooth(method='lm') +
  labs (title = "Relationship Between 'gross' and 'cast_facebook_likes'",
        x = "Number of Likes for Cast on Faceboook", 
        y = "Gross Earnings in the US Box Office",
        caption = "Source: Kaggle IMDB 5000 Movie Dataset") + 
  theme_classic(base_size = 18) +
  scale_x_log10()

```

-   Examine the relationship between `gross` and `budget`.
-   Budget appears to be a good predictor of gross, since the graph shows a positive correlation. Although when the budget is less than 1 million dollars, it is not a good predictor.

```{r, gross_on_budget}
ggplot(movies) + 
  aes(x=budget, y=gross) + 
  geom_point(size =3) + 
  geom_smooth(method='lm', se=FALSE) +
  scale_x_log10() + 
  labs(title = "Relationship Between 'gross' and 'budget'",
        x = "The Movie's Budget", 
        y = "Gross Earnings in the US Box Office",
        caption = "Source: Kaggle IMDB 5000 Movie Dataset") +
  theme_bw(base_size=18)
```

-   Examine the relationship between `gross` and `rating`.
-   Generally, some genre have much more rating than others,for instance, crime, documentary, and mystery. The others, however, have significantly less reviews. For those with high numbers of reviews, such as Action, Adventure, there exist positive correlation between gross and rating which makes IMDB rating a good predictor for those movies.

```{r, gross_on_rating}
ggplot(movies) + 
  aes(x=rating, y=gross, colour = genre) +
  geom_point(size=2) +
  facet_wrap(~genre) +
  labs(title = 'Relationship Between Rating and Gross Revenue By Genre' , 
       x = NULL,
       y = NULL,
       caption = "Source: Kaggle IMDB 5000 Movie Dataset") +
  theme_bw(base_size=18) +
  theme(legend.position = "none") 
```

## Differences Between the Mean IMDB Ratings for Steven Spielberg and Tim Burton

Null Hypothesis: IMDB mean rating is the same for SS and TB. difference in m =0 Alternative hypothesis: difference in mean is not equal to 0. t-stat: difference between mean divided by standard error p-value

```{r load-movies-data,fig.height=5, fig.width=8 }
movies <- read_csv(here::here("data", "movies.csv"))
#glimpse(movies)

movies_2 <- movies %>% 
  group_by(director) %>% 
  filter(director %in% c("Steven Spielberg","Tim Burton")) %>% 
  summarise(rating,mean_rating = mean(rating), #find datapoints needed for t distribution.
         sd_rating = sd(rating),
         count=n(),
         t_critical =qt(0.975,count-1),
         se_rating = sd(rating)/sqrt(count),
         margin_of_error = t_critical*se_rating, 
         rating_low = mean_rating - margin_of_error, 
         rating_high = mean_rating +margin_of_error,
         highest_rating =max(rating),
         lowest_rating = min(rating))

#movies_mean<-count(movies, mean_rating) 
#movies_mean
movies_2

  ggplot(movies_2, aes(x = mean_rating,y = director))+
  geom_point() +
  geom_errorbar(data = movies_2,aes(xmin = rating_low, xmax=rating_high,height=0.2), size = 1, height = 0.1) +
  labs(x = "Mean IMDB Rating",
       y = NULL,
       title = "Mean IMDB Rating for Spielburg and Burton",
       subtitle="95% confidence interval overlap") +
  geom_rect(aes(xmin=7.27, xmax= 7.33, ymin=0, ymax=3),
            fill = "grey70",
            alpha = 0.5)+
  theme_minimal(base_size = 16) #+
  #geom_text(aes(x= mean_rating, label = labels), vjust=0, nudge_y =0.05,overlap=FALSE) +
  #geom_text(label=mean) 
  #theme_bw()

```

As there is an overlap in the confidence interval, we will now need to use t test for further examination

-   With a p-value of 1%\<5%, we can reject the null hypothesis that Spielburg and Burton have the same mean rating.

-   We should use both the `t.test` command and the `infer` package to simulate from a null distribution, where you assume zero difference between the two.

-   By using the infer package, we can simulate from a null distribution and compute p-values (with get_p\_value()). As seen below, the p-value from the null distribution gives 0.4%\<1%\<5%. Thus, we can infer that the null hypothesis can be rejected.

-   We can conclude that the difference in mean ratings for the two directors is unlikely to be zero.

```{r, fig.height=5, fig.width=8}
t.test(rating~director, data = movies_2) #find t-stat, p-value

#simulating a null world using infer()
set.seed(1234)
ratings_in_null_world <- movies_2 %>%
  specify(rating~director) %>% # we want to look at ratings of directors
  hypothesize(null = "independence") %>%  # hypothesize that the difference is 0
  generate(reps=1000, type="permute") %>% # create a bunch of simulated samples
  calculate(stat="diff in means", order = c("Steven Spielberg","Tim Burton")) # find difference in means of each sample. 

#ratings_in_null_world visualize, with shaded p value
ratings_in_null_world %>% 
  visualize() +
  shade_p_value(obs_stat = 0.64, direction = "two-sided")

# finding p-value of simulated distribution
p_value <- ratings_in_null_world%>%
  get_p_value(obs_stat = .64, direction = "two-sided")
p_value

```
