---
categories:  
- ""    #the front matter should be like the one found in, e.g., blog2.md. It cannot be like the normal Rmd we used
- ""
date: "2022-09-18"
description: Share of Renewable Energy Production in the World # the title that will show up once someone gets to this page
draft: false
image: "" # save picture in \static\img\blogs. Acceptable formats= jpg, jpeg, or png . Your iPhone pics wont work

keywords: ""
slug: project2 # slug is the shorthand URL address... no spaces plz
title: Share of Renewable Energy Production in the World
---

The National Bureau of Economic Research (NBER) has a a very interesting dataset on the adoption of about 200 technologies in more than 150 countries since 1800. This is the [Cross-country Historical Adoption of Technology (CHAT) dataset](https://www.nber.org/research/data/cross-country-historical-adoption-technology).

The following is a description of the variables

| **variable** | **class** | **description**                |
|--------------|-----------|--------------------------------|
| variable     | character | Variable name                  |
| label        | character | Label for variable             |
| iso3c        | character | Country code                   |
| year         | double    | Year                           |
| group        | character | Group (consumption/production) |
| category     | character | Category                       |
| value        | double    | Value (related to label)       |

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=9.25, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(dplyr)
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(wbstats)
library(countrycode)
library(patchwork)
library(gganimate)
library(infer)
```

```{r,load_technology_data}
technology <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-19/technology.csv')

#get all technologies
labels <- technology %>% 
  distinct(variable, label)

# Get country names using 'countrycode' package
technology <- technology %>% 
  filter(iso3c != "XCD") %>% 
  mutate(iso3c = recode(iso3c, "ROM" = "ROU"),
         country = countrycode(iso3c, origin = "iso3c", destination = "country.name"),
         country = case_when(
           iso3c == "ANT" ~ "Netherlands Antilles",
           iso3c == "CSK" ~ "Czechoslovakia",
           iso3c == "XKX" ~ "Kosovo",
           TRUE           ~ country))

#make smaller dataframe on energy
energy <- technology %>% 
  filter(category == "Energy")

# download CO2 per capita from World Bank using {wbstats} package
# https://data.worldbank.org/indicator/EN.ATM.CO2E.PC
co2_percap <- wb_data(country = "countries_only", 
                      indicator = "EN.ATM.CO2E.PC", 
                      start_date = 1970, 
                      end_date = 2022,
                      return_wide=FALSE) %>% 
  filter(!is.na(value)) %>% 
  #drop unwanted variables
  select(-c(unit, obs_status, footnote, last_updated))

# get a list of countries and their characteristics
# we just want to get the region a country is in and its income level
countries <-  wb_cachelist$countries %>% 
  select(iso3c,region,income_level)

```

This is a very rich data set, not just for energy and CO2 data, but for many other technologies. In our case, the emphasis is on data manipulation, rather than making the graphs gorgeous.

Firstly, we produce a graph with the countries with the highest and lowest % contribution of renewables in energy production. This is made up of `elec_hydro`, `elec_solar`, `elec_wind`, and `elec_renew_other`. We use the *patchwork* package to assemble the two charts next to each other.

```{r}
library(patchwork)

en_new <- energy %>%  #filter energy dataset for year + create new variable for all renewable energy
  filter(year == 2019) %>% 
  group_by(country, variable) %>% 
  summarise(count = sum(value)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = "variable", values_from = "count") %>% 
  mutate(renew_en = elec_hydro + elec_solar + elec_wind + elec_renew_other)

en_new[is.na(en_new)] <- 0 #establish renewable energy usage as a percentage of energy as a whole
en_new <- en_new %>% 
mutate(percent = renew_en/ elecprod*100) %>% 
arrange(desc(percent)) %>% 
filter(renew_en != 0, percent != Inf)
en_new


#create plot 1 for highest % of countries with renewable energy
c2p1 <- ggplot(en_new %>% slice_max(order_by = percent, n = 20), 
               aes(x = percent, y = fct_reorder(country, percent))) + 
        geom_col() + 
        labs(title = "Highest and Lowest Percentage of Renewables", x = NULL, y = NULL)
c2p1


#create plot 2 for lowest % of countries with renewable energy
dfsid<- en_new %>%
  slice_min(order_by = percent, n=20)


c2p2 <- ggplot(dfsid, aes(x= percent, y=fct_reorder(country, percent))) + 
        geom_col() + 
        labs(title ='graph2',y = NULL, x = NULL)
c2p2

#combining plots
c2p1 + c2p2

```

Secondly, we produced an animation to explore the relationship between CO2 per capita emissions and the deployment of renewables.

-   As the % of energy generated by renewables goes up, CO2 per capita emissions do seem to go down

```{r}
library(gifski)
library(png)

new_co2_percap <- merge(co2_percap, countries, by="iso3c") #merge all the data into one  dataset
new_co2_percap <- merge(new_co2_percap, en_new, by="country")

library(gapminder)
library(gganimate)
library(av)
library(tibble)
data <- new_co2_percap[,c(1,2,6,7,9,21)] #selecting columnns from above dataset
library(rmarkdown)

#plot and animate the graph
ggplot(data, aes(x=percent, y=value, color=income_level)) + 
  geom_point() + 
  facet_wrap(~income_level, nrow = 2) + 
  labs(title = 'Year:{round(frame_time, 0)}', 
       x='% of Renewable Resources', 
       y='CO2 Present Per Cap') + 
  transition_time(date) + 
  ease_aes('linear') +
  theme(legend.position = "none")

```
